### 2013-03-26 doc/dissertation/document/chapters/parallel_mc/parallel_mc.tex
outlining sections for explanation of asynchronous MSOD algorithm

For Monte Carlo methods, and in particular MCSA, to be viable at the
production scale, scalable parallel implementations are required. In a
general linear algebra context for discrete systems, this has not yet
been achieved. Therefore, we will develop and implement parallel
algorithms for these methods leveraging both the knowledge gained from
the general parallel implementations of Krylov methods and modern
parallel strategies for Monte Carlo as developed by the reactor
physics community. In order to formulate a parallel MCSA algorithm, we
recognize that the algorithm occurs in two stages, an outer iteration
performing Richardson's iteration and applying the correction, and an
inner Monte Carlo solver that is generating the correction via the
adjoint method. The parallel aspects of both these components must be
considered. In this chapter we review particle transport methods for
domain decomposed Monte Carlo and then devise an algorithm based on
these methods for efficient parallel solutions using the adjoint
Neumann-Ulam method and MCSA.

@@

%%---------------------------------------------------------------------------%%
\section{Fully Asynchronous MSOD Neumann-Ulam Algorithm}
\label{sec:asynchronous_algorithm}
We can take much of what was learned from the survey of parallel Monte
Carlo methods for radiation transport and directly apply it to a
parallel formulation of our stochastic linear solvers. Direct analogs
can be derived from these works by noting that the primary difference
between solving a linear system with Monte Carlo methods and fixed
source Monte Carlo transport problems is the content of the Markov
chains that are generated. The transitions represented by these chains
are bound by probabilities and weights and are initiated by the
sampling of a source. In the context of transport problems, those
transitions represent events such as particle scattering and
absorption with probabilities that are determined by physical data in
the form of cross sections. For stochastic matrix inversion, those
transitions represent moving between the equations of the linear
system (and therefore the physical domain which they represent) and
their probabilities are defined by the coefficients of those
equations. Ultimately, we tally the contributions to generate
expectation values in the desired states as we progress through the
chains. Therefore, parallel methods for Monte Carlo radiation
transport can be abstracted and we can use those concepts that apply
to matrix inversion methods as an initial means of developing a
parallel Neumann-Ulam-type solver. Based on the results observed in
the last decade of parallel Monte Carlo development, we can generate
many practical questions that this type of work could answer.

>>>>

For Monte Carlo methods, and in particular MCSA, to be viable at the
production scale, scalable parallel implementations are required. In a
general linear algebra context for discrete systems, this has not yet
been achieved. Therefore, we will develop and implement parallel
algorithms for these methods leveraging both the knowledge gained from
the general parallel implementations of Krylov methods and modern
parallel strategies for Monte Carlo as developed by the reactor
physics community. In order to formulate a parallel MCSA algorithm, we
recognize that the algorithm occurs in two stages, an outer iteration
performing Richardson's iteration and applying the correction, and an
inner Monte Carlo solver that is generating the correction via the
adjoint method. The parallel aspects of both these components must be
considered. In this chapter we briefly review particle transport
methods for domain decomposed Monte Carlo and then devise an algorithm
based on these methods for efficient parallel solutions using the
adjoint Neumann-Ulam method and MCSA.

@@

%%---------------------------------------------------------------------------%%
\section{Fully Asynchronous MSOD Neumann-Ulam Algorithm}
\label{sec:asynchronous_algorithm}
We can take much of what was learned from Brunner and Wagner's
parallel Monte Carlo methods for radiation transport and directly
apply it to a parallel formulation of our stochastic linear
solvers. Direct analogs can be derived from these works by noting that
the primary difference between solving a linear system with Monte
Carlo methods and fixed source Monte Carlo transport problems is the
content of the Markov chains that are generated. The transitions
represented by these chains are bound by probabilities and weights and
are initiated by the sampling of a source. In the context of transport
problems, those transitions represent events such as particle
scattering and absorption with probabilities that are determined by
physical data in the form of cross sections. For stochastic matrix
inversion, those transitions represent moving between the equations of
the linear system (and therefore the physical domain which they
represent) and their probabilities are defined by the coefficients of
those equations. Ultimately, we tally the contributions to generate
expectation values in the desired states as we progress through the
chains. Therefore, parallel methods for Monte Carlo radiation
transport can be abstracted and we can use those concepts that apply
to matrix inversion methods as an initial means of developing a
parallel Neumann-Ulam-type solver.

\subsection{Parallel Random Number Generation}
\label{subsec:parallel_rng}

\subsection{Generating the Transport Domain with Overlap and Preconditioning}
\label{subsec:domain_generation}

\subsection{Generating the Transport Source with Preconditioning}
\label{subsec:source_generation}

\subsection{Transition Processing Kernel}
\label{subsec:transition_processing}

\subsection{Local Domain Transport Kernel}
\label{subsec:local_domain_transport}

\subsection{Domain-to-Domain Communication Kernel}
\label{subsec:domain_to_domain_kernel}

\subsection{Asynchronous Transport Kernel}
\label{subsec:async_transport_kernel}

\subsection{Multiple Set Algorithm}
\label{subsec:multiple_sets_algorithm}
